The Changing Face of Pediatric Pulmonary Exacerbations in CF (NACFC 2023)

### 2023-10-10 Scott Sagel/PEx in CF/nacfc_2023.qmd
pex separate data cleaning file

```{r}
#| include: false
library(readxl)
library(tidyverse)
library(gt)
library(gtsummary)
source("~/GitHub/CF/genotype_class_severity.R")
# Import
chco <- read_excel("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Scott Sagel/PEx in CF/Data_Clean/CHCO.xlsx", na = c("", "NULL"))
cnh <- read_excel("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Scott Sagel/PEx in CF/Data_Clean/CNH.xlsx")
# Format CHCO columns
chco$`P. aeruginosa` <- coalesce(chco$`P. aeruginosa mucoid`, chco$`P. aeruginosa non-mucoid`)
chco <- chco %>%
  rename(
    Race = `Race (White/Black/Other)`, Ethnicity = `Hispanic/Latinx`
  ) %>%
  mutate(
    Genotypes1 = case_when(
      Genotypes1 == "F508del" | Genotypes1 == "delF508" ~ "F508del",
      .default = "Other"
    ),
    Genotypes2 = case_when(
      Genotypes2 == "F508del" | Genotypes2 == "delF508" ~ "F508del",
      .default = "Other"
    ),
    `FEV1 % Predicted at Admit` = suppressWarnings(as.numeric(`FEV1 % Predicted at Admit`))
  ) %>%
  unite(Genotype, Genotypes1, Genotypes2, sep = "/")
# Format CNH to match CHCO
cnh <- cnh %>%
  rename(
    `# Days in Admit` = `Total hospitalization days`,
    `Age at Admit` = `Age at admission`,
    `FEV1 % Predicted at Discharge` = `End of treamtent ppFEV1`,
    `Unique number` = `Study ID`
  ) %>%
  mutate(
    Site = "CNH",
    `Modulator at Time of Admit` = case_when(
      None == "Unchecked" ~ "Yes",
      None == "Checked" ~ "No"
    ),
    MRSA = case_when(
      MRSA == "Unchecked" ~ NA,
      MRSA == "Checked" ~ TRUE
    ),
    `Staph aureus` = case_when(
      MSSA == "Unchecked" ~ NA,
      MSSA == "Checked" ~ TRUE
    ),
    `P. aeruginosa` = case_when(
      `P. aeruginosa` == "Unchecked" ~ NA,
      `P. aeruginosa` == "Checked" ~ TRUE
    ),
    Genotype = case_when(
      `CFTR genotype, groups` == "F508del heterozygous" ~ "F508del/Other",
      `CFTR genotype, groups` == "F508del homozygous" ~ "F508del/F508del",
      `CFTR genotype, groups` == "Other" ~ "Other/Other"
    ),
    `Specific Modulator at Admit` = case_when(
      `Ivacaftor (Kalydeco)` == "Checked" ~ "Other",
      `Ivacaftor/lumacaftor (Orkambi)` == "Checked" ~ "Other",
      `Tezacaftor/ivacaftor (Symdeko)` == "Checked" ~ "Other",
      `Elexacaftor/tezacaftor/ivacaftor (Trikafta)` == "Checked" ~ "Trikafta",
      .default = "None"
    ),
    `FEV1 % Predicted at Admit` = as.numeric(`PEx ppFEV1`)
  )
# Merge
df <- full_join(chco, cnh)
# Format column types
df$`Modulator at Time of Admit` <- factor(df$`Modulator at Time of Admit`,
  levels = c("No", "Yes"), ordered = T
)
df$`Specific Modulator at Admit`[df$`Modulator at Time of Admit` == "No"] <- "None"
df$`Specific Modulator at Admit` <- factor(tolower(df$`Specific Modulator at Admit`),
  levels = c("kalydeco", "none", "orkambi", "symdeko", "trikafta"),
  labels = c("Other", "None", "Other", "Other", "Trikafta")
)
df$Race <- factor(df$Race,
  levels = c(
    "Amer. Indian or Alaskan Naitive", "Asian",
    "Black", "Black or African American", "Black/ AA",
    "Decline to Answer", "Hispanic", "Other",
    "Unknown / Not Reported", "White"
  ),
  labels = c(
    "Other", "Other", "Black", "Black", "Black",
    "Unknown / Not Reported", "Other", "Other",
    "Unknown / Not Reported", "White"
  )
)
df$Ethnicity <- factor(df$Ethnicity,
  levels = c(
    "FALSE", "Hispanic or Latino", "HLS",
    "NOT Hispanic or Latino", "NOT HLS",
    "TRUE", "UNKNOWN"
  ),
  labels = c(
    "NOT Hispanic or Latino", "Hispanic or Latino",
    "Hispanic or Latino", "NOT Hispanic or Latino",
    "NOT Hispanic or Latino", "Hispanic or Latino",
    "Unknown"
  )
)
bugs <- c("P. aeruginosa", "P. aeruginosa mucoid", "P. aeruginosa non-mucoid", "Staph aureus", "MRSA", "S. maltophilia")
df[, bugs] <- lapply(df[, bugs], function(c) {
  c[is.na(c)] <- "No"
  c <- factor(c, levels = c("No", TRUE), labels = c("No", "Yes"), ordered = T)
})
df$`S. maltophilia`[df$Site == "CNH"] <- NA
# Make summary table
summ <- df %>%
  group_by(`Unique number`, Site, Year) %>%
  mutate(`Change in FEV1` = `FEV1 % Predicted at Discharge` - `FEV1 % Predicted at Admit`) %>%
  summarise(
    `# of Hospitalizations per Person` = n(),
    across(c(Sex:Genotype, `Specific Modulator at Admit`), ~ first(na.omit(.x))),
    across(c(`Modulator at Time of Admit`, all_of(bugs)), ~ max(.x, na.rm = T)),
    across(
      all_of(c(
        "Age at Admit", "# Days in Admit", "Change in FEV1",
        "FEV1 % Predicted at Admit", "FEV1 % Predicted at Discharge"
      )),
      ~ mean(.x, na.rm = T)
    ),
    .groups = "drop"
  ) %>%
  mutate(Genotype = factor(Genotype,
    levels = c(
      "F508del/F508del", "F508del/Other",
      "Other/F508del", "Other/Other"
    ),
    labels = c(
      "F508del/F508del", "F508del/Other",
      "F508del/Other", "Other/Other"
    )
  )) %>%
  select(
    `Unique number`, `Site`, Year, `# of Hospitalizations per Person`, `# Days in Admit`, `Age at Admit`,
    `Change in FEV1`, `FEV1 % Predicted at Admit`, `FEV1 % Predicted at Discharge`,
    `Modulator at Time of Admit`, `Specific Modulator at Admit`,
    all_of(bugs), Sex:Genotype
  )
```

>>>>

```{r}
#| include: false
library(tidyverse)
library(gtsummary)
load("/Users/timvigers/Dropbox/Work/Vigers/CF/Scott Sagel/PEx in CF/Data_Cleaned/analysis_dataset.RData")
```

1. FEV1 % predicted at discharge was calculated using the `rspiro` package and GLI-2012 equations.
  - We did not distinguish between NE and SE Asian when capturing race, so all Asian participants were assumed to be NE Asian for these calculations (this was picked randomly).
  - Some participants are under age 3 and/or under 1 meter in height, so the ppFEV1 calculations may not be exactly accurate for these people.

```{r}
# Make summary table
summ <- df %>%
  group_by(sid,redcap_data_access_group,admit_year) %>%
  mutate(`Change in FEV1` = ppfev1_discharge - ppfev1_admit) %>%
  summarise(
    `# of Hospitalizations` = n(),
    across(c(sex:genotype), ~ first(na.omit(.x))),
    across(c(`Modulator at Time of Admit`, all_of(bugs)), ~ max(.x, na.rm = T)),
    across(
      all_of(c(
        "Age at Admit", "# Days in Admit", "Change in FEV1",
        "FEV1 % Predicted at Admit", "FEV1 % Predicted at Discharge"
      )),
      ~ mean(.x, na.rm = T)
    ),
    .groups = "drop"
  ) %>%
  select(
    `Unique number`, `Site`, Year, `# of Hospitalizations per Person`, `# Days in Admit`, `Age at Admit`,
    `Change in FEV1`, `FEV1 % Predicted at Admit`, `FEV1 % Predicted at Discharge`,
    `Modulator at Time of Admit`, `Specific Modulator at Admit`,
    all_of(bugs), Sex:Genotype
  )
```
